
---
title: "Metro Melbourne COVID-19 2023 survey; Census 2016 & 2021 residential relocation"
runningheader: "Metro Melbourne COVID-19 2023 survey; Census 2016 & 2021 residential relocation" # only for pdf output
subtitle: "Data analysis" # only for html output
author: "Dr James Reynolds, Public Transport Research Group (PTRG), Monash University"
date: "`r Sys.Date()`"
output:
  tufte::tufte_html: 
    margin_references: false
    citation_package: natbib
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
  tufte::tufte_book:
    citation_package: natbib
    latex_engine: xelatex
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tufte)
# invalidate cache when the tufte version changes
knitr::opts_chunk$set(cache.extra = packageVersion('tufte'))
options(htmltools.dir.version = FALSE)
```


# 1.0 Introduction
As part of the work program for the upcoming Metro Melbourne COVID-19 2023 (second) survey PTRG is comparing the usual place of residence results from the 2016 and 2021 Australian census. The analysis includes looking at Metro Melbourne as a whole, and also the inner, middle and outer regions individually, with those regions being as defined in the 2020 (first) survey. 

This work note summarises the analysis and findings. The next section discusses the datasets retrieved from the Australian Bureau of Statistics (ABS) Table Builder. Section 3 presents the analysis results


# 2.0 Data 

```{r loadsurveydata, echo=TRUE, include = TRUE} 


#Load survey results dataset

#library(haven)
#revised_dataset_corrections_210722 <- read_sav("data/Mode_shift_from_PT_revised_dataset - grouping - data corrections 210722.sav")

```


# 3.0 Results 


